---
title: "Lab 05 - La Quinta is Spanish for next to Denny's, Pt. 2"
author: "Sophie Boyd"
date: "2-13-26"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox) 
```

```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")
```

### Exercise 1

```{r AK-dennys}
dn_ak <- dennys %>%
  filter(state == "AK")
nrow(dn_ak)
```

There are three Denny's locations in Alaska.


```{r AK-laquinta}
lq_ak <- laquinta %>%
  filter(state == "AK")
nrow(lq_ak)
```

There are two La Quinta locations in Alaska.

### Exercise 2

```{r AK-dn-lq}
nrow(dn_ak) * nrow(lq_ak)
```

We need to calculate 6 distances to cover all possible combinations of Denny's and La Quinta locations in Alaska.

### Exercise 3

```{r dn-lq-join}
dn_lq_ak <- full_join(dn_ak, lq_ak,
  by = "state"
)
dn_lq_ak
```

### Exercise 4

There are six observations in the joined data frame. The names of the variables are address, city, state, zip, longitude, latitude, address, city, zip, each with a version followed by .x and .y (with the exception of state).

### Exercise 5

...

### Exercise 6

...

Add exercise headings as needed.
