---
title: "Lab 05 - La Quinta is Spanish for next to Denny's, Pt. 2"
author: "Sophie Boyd"
date: "2-13-26"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox) 
```

```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")
```

### Exercise 1

```{r AK-dennys}
dn_ak <- dennys %>%
  filter(state == "AK")
nrow(dn_ak)
```

There are three Denny's locations in Alaska.


```{r AK-laquinta}
lq_ak <- laquinta %>%
  filter(state == "AK")
nrow(lq_ak)
```

There are two La Quinta locations in Alaska.

### Exercise 2

```{r AK-dn-lq}
nrow(dn_ak) * nrow(lq_ak)
```

We need to calculate 6 distances to cover all possible combinations of Denny's and La Quinta locations in Alaska.

### Exercise 3

```{r dn-lq-join}
dn_lq_ak <- full_join(dn_ak, lq_ak,
  by = "state"
)
dn_lq_ak
```

### Exercise 4

There are six observations in the joined data frame. The names of the variables are address, city, state, zip, longitude, latitude, address, city, zip, each with a version followed by .x and .y (with the exception of state).

### Exercise 5

To add a new variable to a data frame while keeping the existing variables, we use the mutate function. 

```{r haversine}
haversine <- function(long1, lat1, long2, lat2, round = 3) {
  # convert to radians
  long1 <- long1 * pi / 180
  lat1 <- lat1 * pi / 180
  long2 <- long2 * pi / 180
  lat2 <- lat2 * pi / 180

  R <- 6371 # Earth mean radius in km

  a <- sin((lat2 - lat1) / 2)^2 + cos(lat1) * cos(lat2) * sin((long2 - long1) / 2)^2
  d <- R * 2 * asin(sqrt(a))

  return(round(d, round)) # distance in km
}
```

### Exercise 6

```{r distance}
dn_lq_ak <- dn_lq_ak %>%
  mutate(distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y, round = 3))
```

### Exercise 7

```{r min-distance}
dn_lq_ak <- dn_lq_ak %>%
  group_by(address.x) %>%
  mutate(min_distance = min(distance)) %>%
  ungroup()
```

### Exercise 8

```{r AK-plot}
ggplot(data = dn_lq_ak, aes(x = min_distance)) +
  geom_density(fill = "lightblue", alpha = 0.3) +
  labs(title = "Minimum Distances Between Denny's and La Quinta Locations in Alaska",
       x = "Minimum Distance From La Quinta (in miles)",
       y = "Count",
       ) +
  theme_minimal() +
  scale_fill_viridis_d()
```

One Denny's location in Alaska is especially close to a La Quinta (about 2 miles), while the other 2 Denny's locations are between 5 and 6 miles away from a La Quinta. There are very few values in the distribution, but I would tentatively describe this as left-skewed.

##### Descriptives:

```{r AK-sum-stats}
mean(dn_lq_ak$min_distance)
min(dn_lq_ak$min_distance)
max(dn_lq_ak$min_distance)
```

