---
title: "Lab 05 - La Quinta is Spanish for next to Denny's, Pt. 2"
author: "Sophie Boyd"
date: "2-13-26"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox) 
```

```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")
```

### Exercise 1

```{r AK-dennys}
dn_ak <- dennys %>%
  filter(state == "AK")
nrow(dn_ak)
```

There are three Denny's locations in Alaska.


```{r AK-laquinta}
lq_ak <- laquinta %>%
  filter(state == "AK")
nrow(lq_ak)
```

There are two La Quinta locations in Alaska.

### Exercise 2

```{r AK-dn-lq}
nrow(dn_ak) * nrow(lq_ak)
```

We need to calculate 6 distances to cover all possible combinations of Denny's and La Quinta locations in Alaska.

### Exercise 3

```{r dn-lq-join-AK, warning=FALSE}
dn_lq_ak <- full_join(dn_ak, lq_ak,
  by = "state"
)
dn_lq_ak
```

### Exercise 4

There are six observations in the joined data frame. The names of the variables are address, city, state, zip, longitude, and latitude, each with a version followed by .x and .y (with the exception of state).

### Exercise 5

To add a new variable to a data frame while keeping the existing variables, we use the mutate function. 

```{r haversine}
haversine <- function(long1, lat1, long2, lat2, round = 3) {
  # convert to radians
  long1 <- long1 * pi / 180
  lat1 <- lat1 * pi / 180
  long2 <- long2 * pi / 180
  lat2 <- lat2 * pi / 180

  R <- 6371 # Earth mean radius in km

  a <- sin((lat2 - lat1) / 2)^2 + cos(lat1) * cos(lat2) * sin((long2 - long1) / 2)^2
  d <- R * 2 * asin(sqrt(a))

  return(round(d, round)) # distance in km
}
```

### Exercise 6

```{r distance-AK}
dn_lq_ak <- dn_lq_ak %>%
  mutate(distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y, round = 3))
```

### Exercise 7

```{r min-distance-AK}
dn_lq_ak <- dn_lq_ak %>%
  group_by(address.x) %>%
  mutate(min_distance = min(distance)) %>%
  ungroup()
```

### Exercise 8

```{r AK-plot}
ggplot(data = dn_lq_ak, aes(x = min_distance)) +
  geom_density(fill = "lightblue", alpha = 0.3) +
  labs(title = "Minimum Distances Between Denny's and La Quinta Locations in AK",
       x = "Minimum Distance (in miles)",
       y = "Count",
       ) +
  theme_minimal() 
```

##### Descriptives:

```{r AK-sum-stats}
mean(dn_lq_ak$min_distance)
median(dn_lq_ak$min_distance)
min(dn_lq_ak$min_distance)
max(dn_lq_ak$min_distance)
```
One Denny's location in Alaska is especially close to a La Quinta (about 2 miles), while the other 2 Denny's locations are between 5 and 6 miles away from a La Quinta. There are very few values in the distribution, but I would tentatively describe this as left-skewed.


### Exercise 9

#### Filtering Denny's and La Quinta data frames for NC:

```{r NC-dennys}
dn_nc <- dennys %>%
  filter(state == "NC")
nrow(dn_nc)
```

```{r NC-laquinta}
lq_nc <- laquinta %>%
  filter(state == "NC")
nrow(lq_nc)
```

There are 28 Denny's locations and 12 La Quinta locations in North Carolina.

#### Joining the data frames to get a list of all possible pairings:

```{r dn-lq-join-NC, warning=FALSE}
dn_lq_nc <- full_join(dn_nc, lq_nc,
  by = "state"
)
dn_lq_nc
```

#### Calculating the distances between the pairings:

```{r distance-NC}
dn_lq_nc <- dn_lq_nc %>%
  mutate(distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y, round = 3))
```


#### Calculating the minimum distance between each Denny's and La Quinta location:

```{r min-distance-NC}
dn_lq_nc <- dn_lq_nc %>%
  group_by(address.x) %>%
  mutate(min_distance = min(distance)) %>%
  ungroup()
```


#### Visualizing and summarizing:

```{r NC-plot}
ggplot(data = dn_lq_nc, aes(x = min_distance)) +
  geom_density(fill = "lightblue", alpha = 0.3) +
  labs(title = "Minimum Distances Between Denny's and La Quinta Locations in NC",
       x = "Minimum Distance (in miles)",
       y = "Count",
       ) +
  theme_minimal() 
```

##### Descriptives: 

```{r NC-sum-stats}
mean(dn_lq_nc$min_distance)
median(dn_lq_nc$min_distance)
min(dn_lq_nc$min_distance)
max(dn_lq_nc$min_distance)
```
The distribution of minimum distances from La Quinta is skewed right for North Carolina. The distribution of minimum distances is centered around 50-60 miles, but the highest minimum distance is 187.94 miles. 

### Exercise 10

#### Filtering Denny's and La Quinta data frames for TX:

```{r TX-dennys}
dn_tx <- dennys %>%
  filter(state == "TX")
nrow(dn_tx)
```

```{r TX-laquinta}
lq_tx <- laquinta %>%
  filter(state == "TX")
nrow(lq_tx)
```

There are 200 Denny's locations and 237 La Quinta locations in Texas.

#### Joining the data frames to get a list of all possible pairings:

```{r dn-lq-join-TX, warning=FALSE}
dn_lq_tx <- full_join(dn_tx, lq_tx,
  by = "state"
)
dn_lq_tx
```

#### Calculating the distances between the pairings:

```{r distance-TX}
dn_lq_tx <- dn_lq_tx %>%
  mutate(distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y, round = 3))
```

#### Calculating the minimum distance between each Denny's and La Quinta location:

```{r min-distance-TX}
dn_lq_tx <- dn_lq_tx %>%
  group_by(address.x) %>%
  mutate(min_distance = min(distance)) %>%
  ungroup()
```

#### Visualizing and summarizing:

```{r TX-plot}
ggplot(data = dn_lq_tx, aes(x = min_distance)) +
  geom_density(fill = "lightblue", alpha = 0.3) +
  labs(title = "Minimum Distances Between Denny's and La Quinta Locations in TX",
       x = "Minimum Distance (in miles)",
       y = "Count",
       ) +
  theme_minimal() 
```

##### Descriptives: 

```{r TX-sum-stats}
mean(dn_lq_tx$min_distance)
median(dn_lq_tx$min_distance)
min(dn_lq_tx$min_distance)
max(dn_lq_tx$min_distance)
```
The distribution of minimum distances for Texas is very right-skewed. Most of the minimum distances are quite low (centered around 3-6 miles), but the highest minimum distance is 60.58 miles.

### Exercise 11 (state of choice: Arizona)

#### Filtering Denny's and La Quinta data frames for AZ:

```{r AZ-dennys}
dn_az <- dennys %>%
  filter(state == "AZ")
nrow(dn_az)
```

```{r AZ-laquinta}
lq_az <- laquinta %>%
  filter(state == "AZ")
nrow(lq_az)
```

There are 83 Denny's locations and 18 La Quinta locations in Arizona. 

#### Joining the data frames to get a list of all possible pairings:

```{r dn-lq-join-AZ, warning=FALSE}
dn_lq_az <- full_join(dn_az, lq_az,
  by = "state"
)
dn_lq_az
```

#### Calculating the distances between the pairings:

```{r distance-AZ}
dn_lq_az <- dn_lq_az %>%
  mutate(distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y, round = 3))
```

#### Calculating the minimum distance between each Denny's and La Quinta location:

```{r min-distance-AZ}
dn_lq_az <- dn_lq_az %>%
  group_by(address.x) %>%
  mutate(min_distance = min(distance)) %>%
  ungroup()
```

#### Visualizing and summarizing:

```{r AZ-plot}
ggplot(data = dn_lq_az, aes(x = min_distance)) +
  geom_density(fill = "lightblue", alpha = 0.3) +
  labs(title = "Minimum Distances Between Denny's and La Quinta Locations in AZ",
       x = "Minimum Distance (in miles)",
       y = "Count",
       ) +
  theme_minimal() 
```

##### Descriptives: 

```{r AZ-sum-stats}
mean(dn_lq_az$min_distance)
median(dn_lq_az$min_distance)
min(dn_lq_az$min_distance)
max(dn_lq_az$min_distance)
```
The distribution of minimum distances is also right-skewed for Arizona. The median minimum distance from a Denny's to a La Quinta is 7.95 miles, but the mean minimum distance is 36.94 miles and the highest minimum distance is 242.73 miles. 

### Exercise 12

Among Alaska, North Carolina, Texas, and Arizona, I would say that Mitch Hedberg's joke is most applicable to Texas. Texas has more Denny's and La Quinta locations than the other states, and most of the Denny's locations are very close to a La Quinta, as evidenced by the low mean (5.79 miles) and median (3.37 miles) values in the distribution. 
